### 2015-07-06 beamer_stacky/11_UNC_canonical-stacky/Title.tex
add VZB beamer slides

\begin{document}
\title [The canonical ring of a stacky curve]{The canonical ring of a stacky curve}

%\subtitle{Based on Beamer version 3.06}
\author[David Zureick-Brown and John Voight]{David Zureick-Brown (Emory University)
\\ John Voight (Dartmouth)}

%\institute[Emory University]{Emory University\\ Slides available at \url{http://www.mathcs.emory.edu/~dzb/slides/}}
%\institute{Emory University, Department of Mathematics and Computer Science}
\date[Nov 8, 2014]{Automorphic Forms and Related Topics\\ 
Fall Southeastern Sectional Meeting\\
 University of North Carolina at Greensboro, Greensboro, NC\\
\mbox{}\\
Nov 8, 2014\\
\mbox{}\\
% \tiny{Slides available at \url{http://www.mathcs.emory.edu/~dzb/slides/} }
}

\begin{frame}[plain]
  \titlepage
\end{frame}

%%% Use M-x ibuffer t U to replace

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ZBVoight-UNC-canonical-stacky"
%%% End: