### 2015-07-06 beamer_stacky/11_UNC_canonical-stacky/slides/slide00210.tex

\begin{frame}{Petri's theorem}

    Let $C$ be non-hyperelliptic, non-trigonal, not a plane quintic.

  \begin{theorem}[Enriques-Noether-Baggage-Petri]
  The canonical ring $R_C$ is generated in degree 1 with relations in degree 2.
  \end{theorem}

  \begin{remark}
    \begin{enumerate}
    \item For $C$ trigonal or a plane quintic $R_C$ is generated in degree 1 with relations in degrees 2 and 3
    \item (unless $g(C) = 3$, which has a single relation in degree 4)
    \item For $C$ hyperelliptic, there are generators in degrees 1,2, relations in degrees up to 4.
    \end{enumerate}
  \end{remark}

\end{frame}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../ZBVoight-UNC-canonical-stacky"
%%% End: