### 2015-07-06 beamer_stacky/11_UNC_canonical-stacky/slides/slide00220.tex

\begin{frame}{Log Petri's theorem}

    Let $C$ be a curve and $\Delta$ a log divisor.

  \begin{theorem}[Voight, ZB]
  The log canonical ring $R_C$ is generated in degree at most 3 with relations in degree at most 6.
  \end{theorem}

  \begin{remark}
    Lots of exceptional cases if $0 < \deg \Delta  \leq 2$.
  \end{remark}

  \begin{remark}[Things stabilize]
    \begin{enumerate}
    \item Generators in degree 1 with relations in degree 2,3 if
      $\Delta = 3$
    \item   
     (Mumford.) Generators in degree 1 with relations in degree 2 if $\Delta \geq 4$
   \end{enumerate}
 \end{remark}

\end{frame}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../ZBVoight-UNC-canonical-stacky"
%%% End: