### 2015-07-06 beamer_stacky/11_UNC_canonical-stacky/slides/slide00330.tex

\begin{frame}{Main theorem}

  \begin{theorem}[Voight,ZB]
Let $(\XX,\Delta)$ be a tame log stacky curve with signature $(g;e_1,\ldots,e_r;\delta)$ over a field $k$, and let $e=\max(1,e_1,\dots,e_r)$.  Then the canonical ring 
\[ R(\XX,\Delta) = \bigoplus_{d=0}^{\infty} H^0(\XX,\Omega(\Delta)^{\otimes d}) \] 
is generated as a $k$-algebra by elements of degree at most $3e$ with relations of degree at most $6e$.

  \end{theorem}

  \begin{remark}
    Moreover, if $2g-2+\delta \geq 0$, then $R(\XX,\Delta)$ is generated in degree at most $\max(3,e)$ with relations in degree at most $2\max(3,e)$.
  \end{remark}

\end{frame}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../ZBVoight-UNC-canonical-stacky"
%%% End: