### 2015-07-07 beamer_stacky_curves/11_UNC_canonical-stacky/slides.tex
rename folders, import surfaces files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  Modular forms
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \input{slides/slide00100.tex}     %% {Modular forms}
     \input{slides/slide00101.tex}     %% {Ring of Modular forms}
     \input{slides/slide00102.tex}     %% {Ring of Modular forms}
   \input{slides/slide00120.tex}     %% {Rustom's conjectures}
   \input{slides/slide00110.tex}     %% {Main theorem}
   \input{slides/slide00130.tex}     %% {Kodaira--Spencer}
  \input{slides/slide00140.tex}     %% {Example: $X_0(11)$ (fundamental domain))}
   \input{slides/slide00230.tex}     %% {Example of $X_0(11)$}

% \textbf{new slide -- picture for $X_0(11)$.}
% \textbf{new slide -- picture for $X(1)$.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Classical Geometry of curves
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \input{slides/slide00200.tex}     %% {Canonical map/ring}
     % \input{slides/slide00201.tex}     %% {log Canonical map/ring}
     % \input{slides/slide00130.tex}     %% {Kodaira--Spencer} (slide from before)
   \input{slides/slide00210.tex}     %% {Petri's theorem}
   \input{slides/slide00220.tex}     %% {Log petri theorem}
   \input{slides/slide00221.tex}     %% {Log petri theorem, corollary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Stacky curve
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \input{slides/slide00130.tex}     %% {Kodaira--Spencer}
   \input{slides/slide00240.tex}     %% {Fundamental Domain for X(1)}
     \input{slides/slide00241.tex}     %% {Example to show they don't agree} 
   \input{slides/slide00300.tex}     %% {Fractional divisors}
   \input{slides/slide00310.tex}     %% {Floors}
   \input{slides/slide00320.tex}     %% {Example: X(1)}
   \input{slides/slide00330.tex}     %% {Main theorem}
   \input{slides/slide00340.tex}     %% {Final comments}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ZBVoight-UNC-canonical-stacky"
%%% End: