### 2015-07-07 beamer_stacky_curves/11_UNC_canonical-stacky/slides/slide00120.tex

\begin{frame}{Rustom's conjectures (2012)}

  \begin{conjecture}[Rustom]
The $\C$-algebra $M(\Gamma_0(N))$ is generated in weight at most 6 with relations in weight at most 12.    
  \end{conjecture}

-- This was proved by Wagreich in 1980/81.

\vspace{6pt}

  \begin{conjecture}[Rustom]
The $\Z\left[1/6N\right]$-algebra $M(\Gamma_0(N), \Z\left[1/6N\right])$ is generated in weight at most 6 with relations in weight at most 12.    
  \end{conjecture}

-- $M_k(\Gamma_0(N),R)$ consists of forms with $q$-expansion in $R\llbracket q\rrbracket$.

\end{frame}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../ZBVoight-UNC-canonical-stacky"
%%% End: