### 2015-07-07 beamer-curves/11_UNC_canonical-stacky/slides/slide00200.tex

\begin{frame}{Log canonical map/ring}

  \begin{definition}
    The \textbf{canonical map} $\phi_K\colon C \to \P^{g-1}$ is given by $P
    \mapsto [\omega_1(P):\ldots:\omega_g(P)]$.
  \end{definition}

(An embedding iff $C$ is not hyperelliptic.)

  \begin{block}{Facts}

\[C \cong \Proj R_{X,\Delta} \cong  \Proj  \bigoplus_k H^0(X, \Omega^1(\Delta)^{\otimes k}) \]

  \end{block}
  \begin{block}{Facts}
The relations among $R_{X,1}$ are the defining equations of $\phi_K(C)$.
  \end{block}

\end{frame}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../ZBVoight-UNC-canonical-stacky"
%%% End: