### 2016-01-01 curves-write-up/AIF/changes-summary.tex
curves: add some response to e_i -> s_i suggestion

%        File: changes-summary.tex
%     Created: Sat Oct 17 11:00 AM 2015 E
% Last Change: Sat Oct 17 11:00 AM 2015 E
%
%%%%%%%%%%%%%%%%%%%%%
%   AMS packages    %
%%%%%%%%%%%%%%%%%%%%%
\documentclass[10 pt]{amsart}

\usepackage{amsmath}
\usepackage{amsxtra}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{mathrsfs}
\usepackage{subfiles}
\usepackage{mathpazo}
\usepackage{euler}
\usepackage[colorinlistoftodos, textsize=tiny]{todonotes}

\addtolength{\oddsidemargin}{-.5 in}
\addtolength{\evensidemargin}{-.4 in}
\addtolength{\textwidth}{1 in}
\addtolength{\topmargin}{0 in}
\addtolength{\textheight}{0 in}

\RequirePackage{color}
\definecolor{myred}{rgb}{0.75,0,0}
\definecolor{mygreen}{rgb}{0,0.5,0}
\definecolor{myblue}{rgb}{0,0,0.65}

\usepackage{hyperref}
\hypersetup{colorlinks=true,citecolor=blue}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows,decorations.pathmorphing}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{counterexample}[theorem]{Counterexample}
\newtheorem{convention}[theorem]{Convention}
\newtheorem{question}[theorem]{Question}
\newtheorem{conjecture}[theorem]{Conjecture} 
\newtheorem{warning}[theorem]{Warning}
\newtheorem{fact}[theorem]{Fact}
\theoremstyle{remark}
\newtheorem{notation}[theorem]{Notation}
\numberwithin{equation}{section}

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\se{\section}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bn{{\mathbb N}}
\newcommand\bc{{\mathbb C}}
\newcommand\br{{\mathbb R}}
\newcommand\bq{{\mathbb Q}}
\newcommand\bp{{\mathbb P}}
\newcommand\CF{{\mathcal F}}
\newcommand\bz{{\mathbb Z}}
\newcommand\ba{{\mathbb A}}
\newcommand\fa{{\mathfrak a}}
\newcommand\fp{{\mathfrak p}}
\newcommand\fq{{\mathfrak q}}
\newcommand\fm{{\mathfrak m}}
\newcommand\so{{\mathscr O}}
\newcommand\sg{{\mathscr G}}

\newcommand\scm{{\mathscr M}}
\newcommand\scn{{\mathscr N}}
\newcommand\scf{{\mathscr F}}
\newcommand\scg{{\mathscr G}}
\newcommand\sco{{\mathscr O}}
\newcommand\sch{{\mathscr H}}
\newcommand\scl{{\mathscr L}}
\newcommand\sci{{\mathscr I}}

\newcommand \ra{\rightarrow}
\newcommand{\id}{\mathrm{id}}
\newcommand\im{\text{im }}
\newcommand\coker{\text{coker}}
\newcommand \spec{\text{Spec }}
\newcommand \proj{\text{Proj }}
\newcommand \rspec{\textit{Spec }}
\newcommand \rproj{\textit{Proj }}
\newcommand \mg{{\mathscr M_g}}

\DeclareMathOperator\ord{ord}

\newcommand \trdeg{\text{tr. deg }}
\newcommand \codim{\text{codim}}
\newcommand \rk{\text{rk }}
\newcommand \di{\text{div }}
\newcommand \depth{\text{depth }}
\DeclareMathOperator\pic{Pic}
\DeclareMathOperator\lcm{lcm}
\DeclareMathOperator\rank{rank}
\DeclareMathOperator\vol{Vol}
\DeclareMathOperator\supp{Supp}

\def\listtodoname{List of Todos}
\def\listoftodos{\@starttoc{tdo}\listtodoname}

\title{Summary of Changes in response to Referee's Report}
\author{Aaron Landesman, Peter Ruhm, and Robin Zhang}

\begin{document}

\maketitle
\section{Changes Made}
Here are the changes we have made in response to the referee's suggestions.

\begin{enumerate}
	\item We added a remark on Gorensteinness of the log canonical ring following
	the definition of a log spin canonical ring in the background section.
	We noticed in comment 1 that the referee made a comment saying
	``Cohen-Macaulay and
projectively Gorenstein follows as usual (of course, provided there are
no log canonical points).'' We were a bit confused about the parenthetical, but interpreted
it to mean that the canonical ring may not be Gorenstein if it is of the form
\begin{align*}
	\oplus_n H^0(X, P + n L)
\end{align*}
for $L$ some fractional divisor and $P$ a point. Of course, in our paper, all canonical rings are of the form
\begin{align*}
	\oplus_n H^0(X, n L)
\end{align*}
for $L$ a fractional divisor, so this would not pose an issue.
	\item Use Hilbert polynomial in Example $\_\_\_$
	\item Replace ``best lower approximation'' with Hirzebruch--Jung
		fractions
\end{enumerate}

\section{Other suggestions}
The referee also made the following suggestions.

\begin{enumerate}
	\item Change notation from $\frac{e_i - 1}{2 e_i}$
		to $\frac{s_i}{2 s_i + 1}$ where $s_i = \frac{e_i - 1}{2}$ and
		we range over all integers $s_i$ instead of odd integers $e_i$.
\end{enumerate}

We decided not to implement the above suggestions for the following
reasons, respectively.

\begin{enumerate}
	\item Firstly, we maintained the $e_i$ notation for consistency
	with the notation established in Voight--Zureick-Brown. These
	$e_i$ correspond to the stabilizer orders in the log spin
	canonical ring case considered there. Furthermore, we often
	list degrees of generation and relations enumerated by signatures
	of the form $(g; e_1, \ldots, e_n; \delta)$ with bounds given in
	the	form $\min(5, e_1, \ldots, e_n)$ which are more natural to
	consider than enumeration by signatures $(g; s_1, \ldots, s_n;
	\delta)$ with bounds given in the form $\min(5, 2s_1 + 1, \ldots,
	2s_n + 1)$.
\end{enumerate}

\end{document}