### 2017-11-25 surfaces-write-up/MRL/MRL-surfaces-responses-to-referee-3.tex
surfaces: revisions on Aaron's edits + finishing touches

%        File: changes-summary.tex
%     Created: Sat Oct 17 11:00 AM 2015 E
% Last Change: Sat Oct 17 11:00 AM 2015 E
%
%%%%%%%%%%%%%%%%%%%%%
%   AMS packages    %
%%%%%%%%%%%%%%%%%%%%%
\documentclass[10 pt]{amsart}

\usepackage{amsmath}
\usepackage{amsxtra}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{mathrsfs}
\usepackage{subfiles}
\usepackage{mathpazo}
\usepackage{euler}
\usepackage[colorinlistoftodos, textsize=tiny]{todonotes}

\addtolength{\oddsidemargin}{-.5 in}
\addtolength{\evensidemargin}{-.4 in}
\addtolength{\textwidth}{1 in}
\addtolength{\topmargin}{0 in}
\addtolength{\textheight}{0 in}

\RequirePackage{color}
\definecolor{myred}{rgb}{0.75,0,0}
\definecolor{mygreen}{rgb}{0,0.5,0}
\definecolor{myblue}{rgb}{0,0,0.65}

\usepackage{hyperref}
\hypersetup{colorlinks=true,citecolor=blue}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows,decorations.pathmorphing}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{counterexample}[theorem]{Counterexample}
\newtheorem{convention}[theorem]{Convention}
\newtheorem{question}[theorem]{Question}
\newtheorem{conjecture}[theorem]{Conjecture} 
\newtheorem{warning}[theorem]{Warning}
\newtheorem{fact}[theorem]{Fact}
\theoremstyle{remark}
\newtheorem{notation}[theorem]{Notation}
\numberwithin{equation}{section}

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\se{\section}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bn{{\mathbb N}}
\newcommand\bc{{\mathbb C}}
\newcommand\br{{\mathbb R}}
\newcommand\bq{{\mathbb Q}}
\newcommand\bp{{\mathbb P}}
\newcommand\CF{{\mathcal F}}
\newcommand\bz{{\mathbb Z}}
\newcommand\ba{{\mathbb A}}
\newcommand\fa{{\mathfrak a}}
\newcommand\fp{{\mathfrak p}}
\newcommand\fq{{\mathfrak q}}
\newcommand\fm{{\mathfrak m}}
\newcommand\so{{\mathscr O}}
\newcommand\sg{{\mathscr G}}

\newcommand\scm{{\mathscr M}}
\newcommand\scn{{\mathscr N}}
\newcommand\scf{{\mathscr F}}
\newcommand\scg{{\mathscr G}}
\newcommand\sco{{\mathscr O}}
\newcommand\sch{{\mathscr H}}
\newcommand\scl{{\mathscr L}}
\newcommand\sci{{\mathscr I}}

\newcommand \ra{\rightarrow}
\newcommand{\id}{\mathrm{id}}
\newcommand\im{\text{im }}
\newcommand\coker{\text{coker}}
\newcommand \spec{\text{Spec }}
\newcommand \proj{\text{Proj }}
\newcommand \rspec{\textit{Spec }}
\newcommand \rproj{\textit{Proj }}
\newcommand \mg{{\mathscr M_g}}

\DeclareMathOperator\ord{ord}

\newcommand \trdeg{\text{tr. deg }}
\newcommand \codim{\text{codim}}
\newcommand \rk{\text{rk }}
\newcommand \di{\text{div }}
\newcommand \depth{\text{depth }}
\DeclareMathOperator\pic{Pic}
\DeclareMathOperator\lcm{lcm}
\DeclareMathOperator\rank{rank}
\DeclareMathOperator\vol{Vol}
\DeclareMathOperator\supp{Supp}

\usepackage{xr}
\externaldocument{../surface-paper}

\def\listtodoname{List of Todos}
\def\listoftodos{\@starttoc{tdo}\listtodoname}

\title{Summary of Changes in response to Referee's Report}
\author{Aaron Landesman, Peter Ruhm, and Robin Zhang}

\begin{document}

\maketitle
\section{Changes Made}
The authors would like to express their gratitude for
the numerous helpful suggestions given by the referee.
We applied nearly all such suggestions and give here
some brief comments on how we applied certain suggested
changes.
\begin{enumerate}
	\item[(2)] 
		We have added in a detailed example 
		(Example~\autoref{example:effective-example})
		explaining how to apply our method of proof to actually work out
		the generators and relations for a section ring
		associated to an effective divisor on $\bp^m$.

	\item[(10)]
		Response to the concern about base change along inseparable
		extensions:
		We added Remark~\autoref{rem:algebraically-closed}
		explaining why purely inseparable extensions pose
		no additional issue. The point is that cohomology commutes with
		flat base change, and extensions of fields
		are always flat, so the dimensions of the graded pieces of the
		ring will be preserved. Therefore generators
		and relations are preserved under arbitrary base field
		extension. Of course, canonical divisors of root stacks may not
		behave well with respect to base change (as we learned from a
		version of Voight--Zureick-Brown's forthcoming book available
		on Voight's website)
		which is why inseparable extensions
		may cause problems for canonical rings.
		But, for the purposes of this article, we fix a divisor on a
		scheme over some base field, and the structure of that particular section
		ring is unchanged upon base change to the algebraic closure.

	\item[(14) \& (18)]
		We explained in both the projective space case and Hirzebruch case
		that this assumption was nonessential, adding
		Corollary~\autoref{cor:proj-generators-relations} for the projective case
		and Remark~\autoref{rem:hirz-generators-relations-hypothesis} for the
		Hirzebruch case. You are correct that we could have easily changed
		coordinates and reordered the $f_i$ so that the first $m+1$
		many $f_i$'s were $x_0, \ldots, x_m$, but
		we may as well state it with arbitrary independent polynomials so
		as not to require such a change of basis.

	\item[(22)]
		We moved the notation to its own section after the introduction,
		as most of the notation was not necessary to state
		our main theorems.

	\item[(24)]
		Indeed, the proof of \autoref{prop:pm-span-and-basis}	was poorly
		phrased, and we have tried to clarify this now.
		The point is that we are only varying $f_0, \ldots, f_m,$ so
		the resulting monomials in the $f_i$ will be independent because
		products of the $f_i$ are. See the updated proof in the paper for the details.

	\item[(25)]
		It is not an ordering, only a partial ordering (which we have now stated).

	\item[(27)]
		Good point, we should have done a reduction instead of claiming it was analogous.
		This is now included at the beginning of the proof of the theorem
		(now numbered \autoref{thm:proj-one-point}).

	\item[(28)]
		Wouldn't there be something like $\binom{m + c_1- 1}{c_1}$ many
		elements? For example, if	we take $\alpha = 1$, and $H = V(x_0)$
		we will have generators $x_1/x_0, \ldots, x_m/x_0$.
		In any case, we decided to remove the paragraph mentioned by
		this comment as well as the claim about minimality of the
		presentation given in the theorem.
\end{enumerate}
\end{document}